---
title: "p8105_hw2_dd2948"
author: "David DeStephano"
date: "September 28, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE}
library(tidyverse)
```

##Question 1
#Mr Trash Wheel
```{r}
trash<-readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1)

trash<-trash%>%
  janitor::clean_names() %>%
    select(-x15) %>%
  drop_na(dumpster) %>%
  mutate(sports_balls=as.integer(sports_balls))
  
```

#Precipitation
```{r}
precip2017<-readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 6, range = "A2:B14") %>%  
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year=2017) 
  
precip2018<-readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5, range = "A2:B14") %>%  
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year=2018) 

precip2017_2018 <-bind_rows(precip2017, precip2018)%>%
  mutate(month=month.name[month])

```
#############
#Write a Paragraph
#############

It might be obvious, but there are 24 rows in the stacked 2017 and 2018 precipitation data. It's good to knwo there are no missing months! This data is helpful and allows us to think about the Mr. Trashwheel data better. Perhaps months with more precipitations collected more trash. The key variable here is "total" (for total precipitation), and year and month.
``

The Trashwheel dataset includes several key variables with dimensions: `r dim(trash)` (344 observations, 16 variables). Month, year, and all the trash cariables are key variables in this dataset.

Total precipitation in 2018 is `r sum(precip2018$total)` inches.
The median number of sports bals collected in 2017 was 
```{r} 
trash %>% group_by(year) %>%  summarize(median(sports_balls, na.rm=TRUE))
``` 
(The median is 8)




##Question 2
```{r}
pols<-read_csv("./data/fivethirtyeight_datasets/pols-month.csv")

unemployment<-read_csv("./data/fivethirtyeight_datasets/unemployment.csv")

snp<-read_csv("./data/fivethirtyeight_datasets/snp.csv")


pols_tidy <- pols %>%
  separate(mon, into = c("year", "month", "day"), convert=TRUE) %>%
  mutate(month=month.abb[month],
         president=ifelse(prez_gop, "gop", "dem"))%>%
  select(-c(day, prez_gop, prez_dem))%>%
  arrange(year, month)
  
snp_tidy <- snp %>%
  separate(date, into = c("month", "day", "year"), convert=TRUE) %>%
  select(-c(day))%>%
  select(year, everything())%>%
  mutate(month=month.abb[month])%>%
  arrange(year, month)

unemployment_tidy<-unemployment%>%
  pivot_longer(
    Jan:Dec,
    names_to = "month", 
    values_to = "percent_unemployed")%>%
  rename(year=Year)
  

combined_inc<-left_join(pols_tidy, snp_tidy, by=c("year", "month"))

combined<-left_join(combined_inc, unemployment_tidy, by=c("year", "month"))
```  
pols contained a dummy variable for whether the prsident at a given day, month, and year was a democrat, and another for republican, including a date variable.

The new pols_tidy dataset includes one variable for either democrat or republican, and two variables for month and year, instead of one variable that includes the full date.
The new dimensions of pols_tidy is `dim(pols_tidy)` (822 rows, 9 columns). The records started in April 1947 and end in April 2015

The snp data is the S&P points for certain dates, starting in April 1950 and ending inFebruary 2015. There are 787 rows and 3 columns in the new dataset, before there were only 2 columns for S&P at close and the date

The unemployment set contains unemployment rates for dates from Jan 1948 to Dec 2015. The original set included one variable for each month, with one variable for year (each row was one year). The new data set just has three columns, month, year, and the unemployment rate, with 816 observations.

The key variables in the new data set are "president", "close", and "percent_unemployed"

##Problem 3

```{r}
baby_names<-read_csv("./data/Popular_Baby_Names.csv")
#The following table shows the case and naming issues with the Ethnicity variable
baby_names %>% count(Ethnicity) %>%  knitr::kable()

#recode ethnicity variabe, make variables lowercase, make cell contents lowercase to remove since different name cases, duplictates with distinct()
baby_names<- baby_names %>%
  mutate(Ethnicity=recode(Ethnicity, "ASIAN AND PACI"="ASIAN AND PACIFIC ISLANDER", 
                          "BLACK NON HISP"="BLACK NON HISPANIC", 
                          "WHITE NON HISP"="WHITE NON HISPANIC"))%>%
  janitor::clean_names()%>%
  mutate(childs_first_name = str_to_lower(childs_first_name))%>%
  distinct()


baby_names %>% count(ethnicity) %>%  knitr::kable()


#Produce a well-structured, reader-friendly table showing the rank in popularity of the name "Olivia" as a female baby name over time; this should have rows for ethnicities and columns for year. 
olivia<-baby_names%>%
  filter(childs_first_name=="olivia")%>%
  select(-c(count))%>%
  pivot_wider(names_from = "year_of_birth",
              values_from=rank)
View(olivia)
head(olivia)
```

#This was the first method for creating a "similar" table for males. Where the mean rank was taken across all years and ethnicities to find the most popular name oveall. Alternative table (as specified in discussion board) is below

```{r}
#Find the most popular male baby name across all ethnicities
most_popular<-baby_names%>%
  filter(gender=="MALE")%>%
  group_by(childs_first_name)%>%
  summarise(mean = mean(rank, na.rm=TRUE))%>%
  arrange(mean)%>%
  View()
#Aparrently its Moshe?? Ethan is the second best and includes all ethnicities, "Moshe" is only for white people..

#Produce a similar table showing the most popular name among male children over time. Including "Moshe"" just because it has the lowest average, even though there is no nonwhite rank data
popular_males <- baby_names %>%
  filter(gender=="MALE")%>%
  select(-c(count))%>%
  pivot_wider(names_from = "year_of_birth",
              values_from=rank)%>%
  filter(childs_first_name=="moshe" | childs_first_name=="ethan")%>%
  arrange(childs_first_name)%>%
  View()

```

#########################################
#"The second one. You need to create a data frame that has rows for ethnicities and columns for years, and each cell contains the most popular boy name for that ethnicity / year."
#########################################
```{r}
name_as_cell <- baby_names %>%
  filter(gender=="MALE")%>%
  select(-c(count))%>%
  filter(rank==1) %>% 
  pivot_wider(names_from = "year_of_birth",
              values_from=childs_first_name)

View(name_as_cell)
head(name_as_cell)
```

#Scatter plot
Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}
scatter_data<-baby_names %>% filter(gender=="MALE", ethnicity=="WHITE NON HISPANIC", year_of_birth==2016)

scatter_data %>% ggplot(aes(x=rank, y=count)) +
  geom_point() +
  labs(
    title = "Rank by count of names, nonhispanic white males in 2016",
    x = "Rank",
    y = "Count"
  )


```


%>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% %>% 


  
  
  